---
title: "A filtering pipeline for enhancers and promoters"
output: html_document
---

## Prepare input files in one workwing directory
```{r}
# setwd("/flush1/ngu121/Unified_ENHANCERS/FinalConcensusEnhancers/FINAL_88Tis5dat_ExactMuli080/CombinedMatrixFilter/")
# setwd("/datastore/ngu121/HPRSforPig/CombineMatrix/")

##Use the following argument input if run by linux command line###############
# argv <- commandArgs(TRUE)                                                  #
# filename =argv[1]                                                          #
# print(filename)                                                            #
# if (length(argv) < 1) {stop("Missing argument: filename & filename_read")}##

path<-'/datastore/ngu121/HPRSforPig/CombineMatrix/'
name1_Meta <-'susScr3_7filter_metadata_forFiltering.bed'
name2_VillarEn <-'susScr-H3K27Ac_NOToverlapH3K4.bed'
name3_VillarPro <- 'susScr-H3K4me3_replicated-peaks_macs.4bed'

#Index of regions containing Villar Promoter and Enhancer ref
name4_VillarIndex <-'Merged_88Tis5datsuScr3IDs_cbVillarPromoterEnhancer.bed' #this is intersect of predicted regulatory regions with the combined reference Villar promoters and enhancers
name5_VillarEnhIndex <-'Merged_88Tis5datsuScr3IDs_OverlapVillarEnh.bed' #this is intersect of predicted regulatory regions with reference Villar enhancers

datforFilter <-read.table(paste0(path,name1_Meta),sep="\t", header=T ) #main table input for the filter
totalEnh <-length(readLines(paste0(path,name2_VillarEn))) #get total number Villar reference enhancer
totalPro <-length(readLines(paste0(path,name3_VillarPro))) #get total number Villar reference promoter
villarIndexPromEnhDat <-read.table(paste0(path,name4_VillarIndex), sep="\t") #reference table for enhancer and promoter
villarIndexEnhDat <-read.table(paste0(path,name5_VillarEnhIndex), sep="\t") #reference table for enhancer only

```

#Calculate baseline
```{r}

#Unit length = Mb
lengthUniversal <-sum(datforFilter$End-datforFilter$Start+1)/1e6 
basePromUniversal <-totalPro/lengthUniversal 
baseEnhUniversal <-totalEnh/lengthUniversal 
basePromGenome <-totalPro/2808.5
baseEnhGenome <-totalEnh/2808.5 

SummarybaselineU <-list(c("enh_Unfilteredbaseline", baseEnhUniversal),c("prom_Unfilteredbaseline", basePromUniversal),
                        c("enh_Genomebaseline", baseEnhGenome),c("prom_Genomebaseline", basePromGenome))

villarIndexAll <-which(datforFilter$IDs%in%villarIndexPromEnhDat$V4)
villarIndexEnh <-which(datforFilter$IDs%in%villarIndexEnhDat$V4)

RegOverlapVillar <-datforFilter[villarIndexAll,]
RegOverlapVillarEnh <-datforFilter[villarIndexEnh,]

```

## Universal function to calculate coverage after each filter
```{r}
bedtools_sv <-function(stringcmd="bedtools intersect", bed1){
  #create temp file
  a.file=tempfile()
  out1=tempfile()
  out2=tempfile()
  #write into bedformat
  write.table(bed1, a.file,col.names = F, quote=F, row.names=F, sep="\t")
  #create command
  cmdEnh <-paste(stringcmd, "-b", a.file, "-a",paste0(path,name2_VillarEn), "-u", ">", out1)
  cmdProm <-paste(stringcmd, "-b", a.file, "-a",paste0(path,name3_VillarPro), "-u", ">", out2)
  #run command
  cat(cmdEnh, "\n"); cat(cmdProm, "\n")
  system(cmdEnh); system(cmdProm)
  #readTable and find overlap
  Enh <-read.table(out1, header=F); Prom <-read.table(out2, header=F)
  totalMB <-sum(bed1$End -bed1$Start +1)/1e6
  RatioE <-dim(Enh)[1]/totalMB; RatioP <-dim(Prom)[1]/totalMB
  OutputSummary <-c("Enh", dim(Enh)[1], RatioE, "Prom", dim(Prom)[1],RatioP, "LengthMB", totalMB)
  unlink(a.file); unlink(out1); unlink(out2)
  return(OutputSummary)
}
```

####Filtering process starts from here####

##1. CAGE## The code will calculate and print out RatioP and RatioE
```{r}
Regulatory_FtCAGE <-subset(datforFilter, CAGEcount >1)
class(datforFilter$MaxRNAseq)
dim(Regulatory_FtCAGE )
#CAGE=1 and had RNAseq signal
Regulatory_FtCAGE_1 <-subset(datforFilter,  CAGEcount ==1)

Regulatory_FtCAGE_1 <-subset(Regulatory_FtCAGE_1,  log2(MaxRNAseq) >= summary(log2(RegOverlapVillar$MaxRNAseq[RegOverlapVillar$MaxRNAseq>0]))[4])

Regulatory_FtCAGE <-rbind(Regulatory_FtCAGE, Regulatory_FtCAGE_1)

Regulatory_RemainCAGE <- subset(datforFilter, CAGEcount <=1)
indexFtCAGE1 <-which( Regulatory_RemainCAGE$IDs%in% Regulatory_FtCAGE_1$IDs) 
Regulatory_RemainCAGE <-Regulatory_RemainCAGE[-indexFtCAGE1,]

datCAGE <-Regulatory_FtCAGE
bed1 <-datCAGE[,c(2,3,4,1)]
OutputSummaryCAGE<-bedtools_sv("bedtools intersect", bed1)
OutputSummaryCAGE; SummarybaselineU;dim(Regulatory_RemainCAGE); dim(Regulatory_FtCAGE); dim(datforFilter)
```

##2. H3K27Ac##

```{r}
head(Regulatory_RemainCAGE)
Regulatory_FtCAGE_FtH3K27Ac <-subset(Regulatory_RemainCAGE[Regulatory_RemainCAGE$MeanH3K27Ac >0,],
                                              log2(MeanH3K27Ac) >=
summary(log2(RegOverlapVillarEnh[RegOverlapVillarEnh$MeanH3K27Ac>0,]$MeanH3K27Ac))[3])

datH3K27Ac <-Regulatory_FtCAGE_FtH3K27Ac
bed1 <-datH3K27Ac[,c(2,3,4,1)]
OutputSummaryH3K27Ac<-bedtools_sv("bedtools intersect", bed1)                      
OutputSummaryH3K27Ac; SummarybaselineU 

Regulatory_RmCAGE_rmH3K27AcIdx <- which(Regulatory_RemainCAGE$IDs %in% Regulatory_FtCAGE_FtH3K27Ac$IDs)

Regulatory_RmCAGE_rmH3K27Ac <-Regulatory_RemainCAGE[-Regulatory_RmCAGE_rmH3K27AcIdx,]

dim(Regulatory_RmCAGE_rmH3K27Ac); dim(Regulatory_FtCAGE_FtH3K27Ac)

```

##3. SVM##
```{r}
#a predicted enhancer has positive score, and score high than a cutoff to be an enhancer

Regulatory_FtCAGE_FtH3K27Ac_FtSVM <-subset(na.omit(Regulatory_RmCAGE_rmH3K27Ac), Length <5000 & SVM>=summary(RegOverlapVillarEnh$SVM)[5]) 

datSVM <-Regulatory_FtCAGE_FtH3K27Ac_FtSVM

bed1 <-datSVM[,c(2,3,4,1)] 

OutputSummarySVM<-bedtools_sv("bedtools intersect", bed1) 

Regulatory_RmCAGE_rmH3K27Ac_rmSVMIdx <- which(Regulatory_RmCAGE_rmH3K27Ac$IDs %in% Regulatory_FtCAGE_FtH3K27Ac_FtSVM$IDs)

Regulatory_RmCAGE_rmH3K27Ac_rmSVM <-Regulatory_RmCAGE_rmH3K27Ac[-Regulatory_RmCAGE_rmH3K27Ac_rmSVMIdx,]

OutputSummarySVM; SummarybaselineU;dim(Regulatory_RmCAGE_rmH3K27Ac_rmSVM); dim(Regulatory_RmCAGE_rmH3K27Ac); dim(Regulatory_FtCAGE_FtH3K27Ac_FtSVM)

```

##4. RNAseq##
```{r}
t <-Regulatory_RmCAGE_rmH3K27Ac_rmSVM[Regulatory_RmCAGE_rmH3K27Ac_rmSVM$MaxRNAseq>0,]

Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq <-subset(t,log2(t$MaxRNAseq) >= summary(log2(RegOverlapVillar$MaxRNAseq[RegOverlapVillar$MaxRNAseq>0]))[5])

datRNAseq <-Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq 

bed1 <-datRNAseq[,c(2,3,4,1)]

OutputSummaryRNAseq<-bedtools_sv("bedtools intersect", bed1)                      

Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseqIndx <-which(Regulatory_RmCAGE_rmH3K27Ac_rmSVM$IDs %in% Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq$IDs)

Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq <-Regulatory_RmCAGE_rmH3K27Ac_rmSVM[-Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseqIndx,]

OutputSummaryRNAseq; SummarybaselineU; dim(Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq); dim(Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq)

```

##5. Annotation per bp##
```{r}
Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot <-subset(Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq , AnnotationCount >= summary(RegOverlapVillar$AnnotationCount)[4]) 

datAnn <-Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot
bed1 <-datAnn[,c(2,3,4,1)] 
dim(bed1)
OutputSummaryAnn<-bedtools_sv("bedtools intersect", bed1)                      

Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnnIdx <- which(Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq$IDs %in% Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot$IDs)

Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnn <-Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq[-Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnnIdx ,]

OutputSummaryAnn; SummarybaselineU;  dim(Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnn); dim(Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot); dim(Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq )  
```

##6. PhastCons##                                       
```{r}
t <-as.numeric(as.vector(RegOverlapVillar$phastCons))
tt <-Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnn

tt$phastCons <-as.numeric(as.vector(tt$phastCons))

Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons <-subset(tt, na.omit(tt$phastCons) >= quantile(na.omit(t),c(.95))) 

datPhastCons<-Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons

bed1 <-datPhastCons[,c(2,3,4,1)] 

OutputSummaryPhastCons<-bedtools_sv("bedtools intersect", bed1)                      
Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnnot_rmPhastConsIdx <-which(tt$IDs %in% Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons$IDs) 

Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnnot_rmPhastCons<-tt[-Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnnot_rmPhastConsIdx,]

OutputSummaryPhastCons; SummarybaselineU; dim(Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnnot_rmPhastCons); dim(tt)

```
##7. TFBS##
```{r}
# Regulatory_RmCAGE_rmRNAseq_rmH3K27Ac_rmSVM_rmAnn[Regulatory_RmCAGE_rmRNAseq_rmH3K27Ac_rmSVM_rmAnn$IDs=="88Tis5dat3Type_ID268243",]

dat <-Regulatory_RmCAGE_rmH3K27Ac_rmSVM_rmRNAseq_rmAnnot_rmPhastCons

Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons_FtcountTFBS <-subset(dat,TFBSCount>= quantile(na.omit(RegOverlapVillar$TFBSCount),c(.95))) 

summary(RegOverlapVillar$TFBSCount)

datTFBS<-Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons_FtcountTFBS

bed1 <-datTFBS[,c(2,3,4,1)]
dim(bed1)
OutputSummaryTFBS<-bedtools_sv("bedtools intersect", bed1)                      

dim(Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons_FtcountTFBS); OutputSummaryTFBS; SummarybaselineU 

```
##Combined filtered
```{r}
Combined7Filters <-rbind(
      Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons_FtcountTFBS,
      Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot_FtPhastCons,
      Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq_FtAnnot,
      Regulatory_FtCAGE_FtH3K27Ac_FtSVM_FtRNAseq,
      Regulatory_FtCAGE_FtH3K27Ac_FtSVM,
      Regulatory_FtCAGE_FtH3K27Ac,
      Regulatory_FtCAGE)

dim(Combined7Filters)
datCombined <-Combined7Filters
bed1 <-datCombined[,c(2,3,4,1)] 
OutputSummaryCombined<-bedtools_sv("bedtools intersect", bed1)                      
OutputSummaryCombined; SummarybaselineU 

```
# Write output and calculate coverage of Villar enhancer and promoter numbers

```{r}

percentGenome <-311.5/2808.5; 
percentEnhancer <-23456/31971
percentPromoter <-11818/13769

write.table(Combined7Filters,"Filtered_7steps_susScr3Sept25_Regulatory.bed", col.names=F, row.names=F, quote=F, sep="\t" )


```













